---
title: "Class 11: Structural Bioinformatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database for biomolecular structure data

> Q1:  Determine the percentage of structures solved by X-Ray and Electron Microscopy.  

Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”

```{r}
data = read.csv("Data Export Summary.csv", header=TRUE)
```
Total Number of Entries:
```{r}
sum(data$Total)
```
Proportion of entries from each method:
```{r}
round((data$Total/sum(data$Total)) * 100, 2)
```

> Also, can you determine what proportion of structures are protein?

```{r}
round(sum(data$Proteins)/sum(data$Total) * 100, 2)
```

## HIV-1 Protease Analysis

> Q2:  Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

3039 results

> Q3:  Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?

We see only one atom per water molecule in the structure because the 

> Q4:  There is a conserved water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have?

> Optional: Generate and save a figure clearly showing the two distinct chains of HIV-protease along with the ligand. You might also consider showing the catalytic residues ASP 25 in each chain (we recommend Licorice for these side-chains).
