---
title: 'Class 9: Unsupervised Learning Mini-Project'
author: "Patrick Paxson"
date: "10/29/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Input

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

Note that the `id` and `diagnosis` columns will not be used for most of the following steps.

we have `r nrow(wisc.df)` samples in this dataset

```{r}
nrow(wisc.df)
```

How many benign (non-cancerous) and malginant (cancerous) samples do we have in the  dataset?

```{r}
table(wisc.df$diagnosis)
```

```{r}
# Convert the features of the data: wisc.data
wisc.data <- as.matrix(wisc.df[,3:32])

# Set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

head(wisc.data)
```

Store the  diagnosis for reference in the future as a separate vector

```{r}
diagnosis <- wisc.df$diagnosis
```

- Q1. How many observations are in this dataset?

```{r}
nrow(wisc.df)
```

- Q2. How many of the observations have a malignant diagnosis?

```{r}
table(wisc.df$diagnosis=="M")
```

- Q3. How many variables/features in the data are suffixed with `_mean`?

```{r}
length(grep("_mean", colnames(wisc.df)))
```

```{r}
# Check column means and standard deviations
colMeans(wisc.data)

apply(wisc.data,2,sd)
```

```{r}
# Perform PCA on wisc.data
wisc.pr <- prcomp(wisc.data, scale=TRUE)

# Look at summary of results
summary(wisc.pr)
```

```{r}
round(colMeans(wisc.data), 3)
```

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis)
```

## Questions


- Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

```{r}
x <- summary(wisc.pr)
x$importance[,1]
```

- Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

```{r}
length(x$importance[3,]>0.7)
```

- Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

```{r}
length(x$importance[3,]>0.9)
```

```{r}
biplot(wisc.pr)
```

```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)

data.dist <- dist(data.scaled)

wisc.hclust <- hclust(data.dist, "single")
```

```{r}
plot(wisc.pr)
abline(wisc.pr, col="red", lty=2)
```

##  Section  7: Prediction

We will use the predict() function that will take our PCA model from before and new cancer cell data and project that data onto our PCA space.

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

Plot theese new samples on our PCA 1 vs PCA 2 plot...

```{r}
grps <- cutree(wisc.hclust, k=2)
g <- as.factor(grps)
g <- relevel(g,2)

plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```